\documentclass[12pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{color}
\usepackage{float}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{tcolorbox}
\tcbuselibrary{listingsutf8}
\graphicspath{{assets/}}

% Header/Footer
\pagestyle{fancy}
\setlength{\headheight}{14.5pt} % Fix fancyhdr warning
\fancyhf{}
\lhead{Contrastive Representation Learning}
\rhead{Lilian Weng}
\cfoot{\thepage}

% Section formatting
\titleformat{\section}
  {\normalfont\Large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}
  {\normalfont\large\bfseries}{\thesubsection}{1em}{}


\hypersetup{
    pdftitle={Contrastive Representation Learning},
    pdfauthor={Lilian Weng},
    pdfkeywords={example, pdf, latex},
    pdfsubject={Example PDF generated from LaTeX},
    pdfcreator={LaTeX with hyperref package},
    pdfproducer={pdflatex},
    pdfstartview={FitH},
    colorlinks=true,
    linkcolor=blue,
    urlcolor=blue
}


\title{Contrastive Representation Learning}
\date{This PDF conversion (v1) was generated on \today\footnotemark}


\begin{document}

\maketitle
\begin{center}
    Online version: \url{https://lilianweng.github.io/posts/2021-05-31-contrastive/} \\
    \textbf{Date:} May 31, 2021 \quad
    \textbf{Estimated Reading Time:}  39 mins \quad
    \textbf{Author:} Lilian Weng
\end{center}

The goal of contrastive representation learning is to learn such an embedding space in which similar sample pairs stay close to each other while dissimilar ones are far apart. Contrastive learning can be applied to both supervised and unsupervised settings. When working with unsupervised data, contrastive learning is one of the most powerful approaches in \href{https://lilianweng.github.io/posts/2019-11-10-self-supervised/}{self-supervised learning}.

\footnotetext{This file was prepared by Sakura (\href{mailto:bili_sakura@zju.edu.cn}{bili\_sakura@zju.edu.cn}). Please contact for any issues or copyright concerns. The original source of this file is \url{https://lilianweng.github.io/posts/2021-05-31-contrastive/}.}

\newpage
\tableofcontents
\newpage
\vspace{1em}
\n\input{content.tex}\n\n\section{Citation}
Please cite this work as:

\begin{tcolorbox}[colback=black!5!white, colframe=black!80!white, boxrule=0.5pt, arc=2mm, left=1mm, right=1mm, top=1mm, bottom=1mm]
  Weng, Lilian. \textit{Contrastive Representation Learning}. Lil'Log (May 2021). \url{https://lilianweng.github.io/posts/2021-05-31-contrastive/}
\end{tcolorbox}

Or use the Bib\TeX{} citation:

\begin{tcolorbox}[
    colback=black!5!white,
    colframe=black!80!white,
    boxrule=0.5pt,
    arc=2mm,
    left=1mm,
    right=1mm,
    top=1mm,
    bottom=1mm,
    listing only,
    listing options={
      basicstyle=\ttfamily\footnotesize,
      breaklines=true,
      literate={\{}{{\char`\{}}1 {\}}{{\char`\}}}1
    }
  ]

@article\{weng2021contrastive,

  title   = \{Contrastive Representation Learning\},

  author  = \{Weng, Lilian.\},

  journal = \{lilianweng.github.io\},

  year    = \{2021\},

  month   = \{May\},

  url     = \{https://lilianweng.github.io/posts/2021-05-31-contrastive/\}

\}
\end{tcolorbox}

% \newpage
\section{References}

\begin{enumerate}
    \item Sumit Chopra, Raia Hadsell and Yann LeCun. \textit{Learning a similarity metric discriminatively, with application to face verification.} CVPR 2005. \href{http://yann.lecun.com/exdb/publis/pdf/chopra-05.pdf}{[pdf]}
    \item Florian Schroff, Dmitry Kalenichenko and James Philbin. \textit{FaceNet: A Unified Embedding for Face Recognition and Clustering.} CVPR 2015. \href{https://arxiv.org/abs/1503.03832}{[arXiv]}
    \item Hyun Oh Song et al. \textit{Deep Metric Learning via Lifted Structured Feature Embedding.} CVPR 2016. \href{https://arxiv.org/abs/1511.06452}{[arXiv]}, \href{https://github.com/rksltnl/Deep-Metric-Learning-CVPR16}{[code]}
    \item Ruslan Salakhutdinov and Geoff Hinton. \textit{Learning a Nonlinear Embedding by Preserving Class Neighbourhood Structure.} AISTATS 2007. \href{http://proceedings.mlr.press/v2/salakhutdinov07a.html}{[pdf]}
    \item Michael Gutmann and Aapo Hyvärinen. \textit{Noise-contrastive estimation: A new estimation principle for unnormalized statistical models.} AISTATS 2010. \href{http://proceedings.mlr.press/v9/gutmann10a.html}{[pdf]}
    \item Kihyuk Sohn et al. \textit{Improved Deep Metric Learning with Multi-class N-pair Loss Objective.} NIPS 2016. \href{https://papers.nips.cc/paper/2016/hash/6b180037abbebea991d8b1232f8a8ca9-Abstract.html}{[pdf]}
    \item Nicholas Frosst, Nicolas Papernot and Geoffrey Hinton. \textit{Analyzing and Improving Representations with the Soft Nearest Neighbor Loss.} ICML 2019. \href{http://proceedings.mlr.press/v97/frosst19a.html}{[pdf]}
    \item Tongzhou Wang and Phillip Isola. \textit{Understanding Contrastive Representation Learning through Alignment and Uniformity on the Hypersphere.} ICML 2020. \href{https://arxiv.org/abs/2005.10242}{[arXiv]}, \href{https://ssnl.github.io/hypersphere/}{[code]}
    \item Zhirong Wu et al. \textit{Unsupervised feature learning via non-parametric instance-level discrimination.} CVPR 2018. \href{https://arxiv.org/abs/1805.01978}{[arXiv]}
    \item Ekin D. Cubuk et al. \textit{AutoAugment: Learning augmentation policies from data.} arXiv preprint arXiv:1805.09501 (2018). \href{https://arxiv.org/abs/1805.09501}{[arXiv]}
    \item Daniel Ho et al. \textit{Population Based Augmentation: Efficient Learning of Augmentation Policy Schedules.} ICML 2019. \href{https://arxiv.org/abs/1905.05393}{[arXiv]}
    \item Ekin D. Cubuk \& Barret Zoph et al. \textit{RandAugment: Practical automated data augmentation with a reduced search space.} arXiv preprint arXiv:1909.13719 (2019). \href{https://arxiv.org/abs/1909.13719}{[arXiv]}
    \item Hongyi Zhang et al. \textit{mixup: Beyond Empirical Risk Minimization.} ICLR 2017. \href{https://arxiv.org/abs/1710.09412}{[arXiv]}
    \item Sangdoo Yun et al. \textit{CutMix: Regularization Strategy to Train Strong Classifiers with Localizable Features.} ICCV 2019. \href{https://arxiv.org/abs/1905.04899}{[arXiv]}
    \item Yannis Kalantidis et al. \textit{Mixing of Contrastive Hard Negatives.} NeuriPS 2020. \href{https://arxiv.org/abs/2010.01028}{[arXiv]}
    \item Ashish Jaiswal et al. \textit{A Survey on Contrastive Self-Supervised Learning.} arXiv preprint arXiv:2011.00362 (2021). \href{https://arxiv.org/abs/2011.00362}{[arXiv]}
    \item Jure Zbontar et al. \textit{Barlow Twins: Self-Supervised Learning via Redundancy Reduction.} arXiv preprint arXiv:2103.03230 (2021). \href{https://arxiv.org/abs/2103.03230}{[arXiv]}, \href{https://github.com/facebookresearch/barlowtwins}{[code]}
    \item Alec Radford, et al. \textit{Learning Transferable Visual Models From Natural Language Supervision.} arXiv preprint arXiv:2103.00020 (2021). \href{https://arxiv.org/abs/2103.00020}{[arXiv]}
    \item Mathilde Caron et al. \textit{Unsupervised Learning of Visual Features by Contrasting Cluster Assignments (SwAV).} NeuriPS 2020. \href{https://arxiv.org/abs/2006.09882}{[arXiv]}
    \item Mathilde Caron et al. \textit{Deep Clustering for Unsupervised Learning of Visual Features.} ECCV 2018. \href{https://arxiv.org/abs/1807.05520}{[arXiv]}
    \item Prannay Khosla et al. \textit{Supervised Contrastive Learning.} NeurIPS 2020. \href{https://arxiv.org/abs/2004.11362}{[arXiv]}
    \item Aaron van den Oord, Yazhe Li \& Oriol Vinyals. \textit{Representation Learning with Contrastive Predictive Coding.} arXiv preprint arXiv:1807.03748 (2018). \href{https://arxiv.org/abs/1807.03748}{[arXiv]}
    \item Jason Wei and Kai Zou. \textit{EDA: Easy data augmentation techniques for boosting performance on text classification tasks.}  EMNLP-IJCNLP 2019. \href{https://arxiv.org/abs/1901.11196}{[arXiv]}
    \item Sosuke Kobayashi. \textit{Contextual Augmentation: Data Augmentation by Words with Paradigmatic Relations.} NAACL 2018. \href{https://arxiv.org/abs/1805.06201}{[arXiv]}
    \item Hongchao Fang et al. \textit{CERT: Contrastive self-supervised learning for language understanding.} arXiv preprint arXiv:2005.12766 (2020). \href{https://arxiv.org/abs/2005.12766}{[arXiv]}
    \item Dinghan Shen et al. \textit{A Simple but Tough-to-Beat Data Augmentation Approach for Natural Language Understanding and Generation.} arXiv preprint arXiv:2009.13818 (2020). \href{https://arxiv.org/abs/2009.13818}{[arXiv]}, \href{https://github.com/dinghanshen/cutoff}{[code]}
    \item Tianyu Gao et al. \textit{SimCSE: Simple Contrastive Learning of Sentence Embeddings.} arXiv preprint arXiv:2104.08821 (2020). \href{https://arxiv.org/abs/2104.08821}{[arXiv]}, \href{https://github.com/princeton-nlp/SimCSE}{[code]}
    \item Nils Reimers and Iryna Gurevych. \textit{Sentence-BERT: Sentence embeddings using Siamese BERT-networks.} EMNLP 2019. \href{https://arxiv.org/abs/1908.10084}{[arXiv]}
    \item Jianlin Su et al. \textit{Whitening sentence representations for better semantics and faster retrieval.} arXiv preprint arXiv:2103.15316 (2021). \href{https://arxiv.org/abs/2103.15316}{[arXiv]}, \href{https://github.com/bojone/BERT-whitening}{[code]}
    \item Yan Zhang et al. \textit{An unsupervised sentence embedding method by mutual information maximization.} EMNLP 2020. \href{https://arxiv.org/abs/2009.12061}{[arXiv]}, \href{https://github.com/yanzhangnlp/IS-BERT}{[code]}
    \item Bohan Li et al. \textit{On the sentence embeddings from pre-trained language models.} EMNLP 2020. \href{https://arxiv.org/abs/2011.05864}{[arXiv]}
    \item Lajanugen Logeswaran and Honglak Lee. \textit{An efficient framework for learning sentence representations.} ICLR 2018. \href{https://arxiv.org/abs/1803.02893}{[arXiv]}
    \item Joshua Robinson, et al. \textit{Contrastive Learning with Hard Negative Samples.} ICLR 2021. \href{https://arxiv.org/abs/2010.04592}{[arXiv]}
    \item Ching-Yao Chuang et al. \textit{Debiased Contrastive Learning.} NeuriPS 2020. \href{https://arxiv.org/abs/2007.00224}{[arXiv]}
\end{enumerate}



\n\end{document}\n